---
title: "Get SOE stock status indicator from assessment data"
author: "Sarah Gaichas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(here)
library(tidyverse)
library(DT)
library(assessmentdata)
library(ggrepel)
library(ecodata)

```

## The old way

I searched for each assessment document and filled in this spreadsheet by hand. Time consuming and error prone. 

```{r 2019assess}

datatable(read.csv(here("2019assess.csv")), rownames = FALSE, options = list(scrollX = TRUE))

```

## The new way

Andy made the [assessmentdata package](https://github.com/NOAA-EDAB/assessmentdata) based on [Stock SMART](https://www.st.nmfs.noaa.gov/stocksmart?app=homepage). Surely this will be better?

Two data frames are in the package, `stockAssessmentData` and `stockAssessmentSummary`.

In `stockAssessmentData` we have time series. Columns are `r names(stockAssessmentData)` and the reported metrics are `r unique(stockAssessmentData$Metric)`. 


```{r}

datatable(head(stockAssessmentData), rownames = FALSE)

```

In `stockAssessmentSummary` we have assessment metadata. Columns are `r (names(stockAssessmentSummary))`.

```{r}

datatable(head(stockAssessmentSummary), rownames = FALSE, options = list(scrollX = TRUE))

```

Trying to rebuild my spreadsheet from `stockAssessmentSummary` and use the ecodata code to make the dataset:

```{r make-2019assess}

test2019assess <- stockAssessmentSummary %>%
  filter(`Science Center` == "NEFSC") %>%
  select(c(`Stock Name`, Jurisdiction, FMP, `Science Center`, 
           `Stock Area`, `Assessment Year`, `Last Data Year`,
           `F Year`, `Estimated F`, Flimit, Fmsy, `F/Flimit`, 
           `F/Fmsy`, Ftarget, `F/Ftarget`, `B Year`, `Estimated B`,
           `B Unit`, Blimit, Bmsy, `B/Blimit`, `B/Bmsy`)) %>%
  arrange(Jurisdiction, `Stock Name`, FMP, `Assessment Year`) %>%
  rename(Entity.Name = `Stock Name`,
         Assessment.Year = `Assessment Year`,
         F.Fmsy = `F/Fmsy`,
         B.Bmsy = `B/Bmsy`)

# from get_stocks.R, ecodata 2020
  #assess <- read.csv(file.path(data.dir, "2019assess.csv"))
  assess <- test2019assess
  #decode <- read.csv(file.path(data.dir, "2019decoder.csv"))
  decode <- read.csv(here("2019decoder.csv"))

stock_status_stockSMART <-
    assess %>%
    dplyr::group_by(Entity.Name) %>%
    dplyr::filter(Assessment.Year == max(Assessment.Year)) %>%
  #Find last year assessment occurred for each stock
    dplyr::ungroup() %>%
    dplyr::left_join(.,decode, by = "Entity.Name") %>% #Join in list of managed species
    dplyr::select(Entity.Name, Assessment.Year, F.Fmsy, B.Bmsy, Council, Code) %>%
  #select column variables to keep
    dplyr::mutate(id = 1:length(Entity.Name)) %>%
    tidyr::gather(.,Var, Value,-id,-Entity.Name,-Assessment.Year,-Council,-Code) %>%
  #wide to long
    dplyr::select(-id) %>%
    dplyr::rename(`Last assessment` = Assessment.Year,
                  Stock = Entity.Name) %>% #rename variables for clarity
    dplyr::mutate(Units = "unitless") %>%
    dplyr::mutate(Value = replace(Value, which(Code == "N Windowpane" & Var == "F.Fmsy"), NA))



```

Then test to see if I get the same plot for assessments up to Dec 2019 I got the old way.

### Comparisons {.tabset}

#### StockSMART source, Mid-Atlantic

```{r, MAFMCplot}

stock_status <- 
  stock_status_stockSMART %>%
  mutate(Code = recode(Code, "Dogfish" = "Sp. Dogfish" )) %>% 
  spread(.,Var,Value) %>% 
  filter(Council %in% c("MAFMC","Both")) %>% 
  group_by(Stock) %>% 
  mutate(score = case_when(
    (B.Bmsy <0.5) ~"a",
    (F.Fmsy >1) ~ "a", 
    (F.Fmsy < 1 & B.Bmsy > 0.5 & B.Bmsy < 1) ~ "b",
    (F.Fmsy < 1 & B.Bmsy > 1) ~ "c"))
#Plot constants
y.max <- 2.0 #1.75 mackerel cut off F/Fmsy is 1.8
x.max <- 2.6
#A dataframe that defines custom legend for stocks with unknown status
unknown <- data.frame(text = c("Unknown Status", "Longfin Squid",
                              "Shortfin Squid", "N. Goosefish", "S. Goosefish"),
                    x = rep(0.9*x.max,5), y = seq(0.93*y.max,1.4,-.1))

# Custom Color
custom_color<- c("#56B4E9", "#009E73", "#0072B2")
#Plotting code
ggplot(data = stock_status) +
  geom_vline(xintercept = 1, linetype = "dotted")+
  geom_vline(xintercept = 0.5, linetype = "dashed")+
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_point(aes(x = B.Bmsy,
                 y = F.Fmsy,
                 shape = Council,
                 color = score)) +
  geom_text_repel(aes(x = B.Bmsy, #geom_text_repel auto-jitters text around points
                      y = F.Fmsy,
                      label = Code, 
                      color = score), 
                  show.legend = FALSE, nudge_y = -0.01, nudge_x = 0.05) +
  scale_color_brewer(palette = "Dark2",
                     breaks = stock_status$score) +
  ylim(0,y.max) +
  xlim(0,x.max) +
  geom_text(data = unknown, aes(x = x, y = y, label = text), #Custom legend for unknown stock status
            size = c(4.75,rep(4,4))) +
  annotate("rect", xmin = 0.8*x.max,
           xmax = x.max,
           ymin = 0.65*y.max,
           ymax = 0.90*y.max,
           alpha = 0.1) +
  xlab(expression(~B/B[msy])) +
  ylab(expression(~F/F[msy])) +
  guides(color = FALSE) +
  theme_ts()
```

#### SOE 2020 ecodata source, Mid-Atlantic

```{r, MAFMC2020plot}
stock_status <- 
  ecodata::stock_status %>%
  mutate(Code = recode(Code, "Dogfish" = "Sp. Dogfish" )) %>% 
  spread(.,Var,Value) %>% 
  filter(Council %in% c("MAFMC","Both")) %>% 
  group_by(Stock) %>% 
  mutate(score = case_when(
    (B.Bmsy <0.5) ~"a",
    (F.Fmsy >1) ~ "a", 
    (F.Fmsy < 1 & B.Bmsy > 0.5 & B.Bmsy < 1) ~ "b",
    (F.Fmsy < 1 & B.Bmsy > 1) ~ "c"))
#Plot constants
y.max <- 2.0 #1.75 mackerel cut off F/Fmsy is 1.8
x.max <- 2.6
#A dataframe that defines custom legend for stocks with unknown status
unknown <- data.frame(text = c("Unknown Status", "Longfin Squid",
                              "Shortfin Squid", "N. Goosefish", "S. Goosefish"),
                    x = rep(0.9*x.max,5), y = seq(0.93*y.max,1.4,-.1))

# Custom Color
custom_color<- c("#56B4E9", "#009E73", "#0072B2")
#Plotting code
ggplot(data = stock_status) +
  geom_vline(xintercept = 1, linetype = "dotted")+
  geom_vline(xintercept = 0.5, linetype = "dashed")+
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_point(aes(x = B.Bmsy,
                 y = F.Fmsy,
                 shape = Council,
                 color = score)) +
  geom_text_repel(aes(x = B.Bmsy, #geom_text_repel auto-jitters text around points
                      y = F.Fmsy,
                      label = Code, 
                      color = score), 
                  show.legend = FALSE, nudge_y = -0.01, nudge_x = 0.05) +
  scale_color_brewer(palette = "Dark2",
                     breaks = stock_status$score) +
  ylim(0,y.max) +
  xlim(0,x.max) +
  geom_text(data = unknown, aes(x = x, y = y, label = text), #Custom legend for unknown stock status
            size = c(4.75,rep(4,4))) +
  annotate("rect", xmin = 0.8*x.max,
           xmax = x.max,
           ymin = 0.65*y.max,
           ymax = 0.90*y.max,
           alpha = 0.1) +
  xlab(expression(~B/B[msy])) +
  ylab(expression(~F/F[msy])) +
  guides(color = FALSE) +
  theme_ts()
```


#### StockSMART source, New England

```{r, NEFMCtest}

stock_status <- stock_status_stockSMART %>%
  mutate(Code = recode(Code, "Dogfish" = "Sp. Dogfish" )) %>% 
  spread(.,Var,Value) %>% 
  filter(Council %in% c("NEFMC","Both")) %>% 
  group_by(Stock) %>% 
  mutate(score = case_when(
    (B.Bmsy <0.5) ~"a",
    (F.Fmsy >1) ~ "a", 
    (F.Fmsy < 1 & B.Bmsy > 0.5 & B.Bmsy < 1) ~ "b",
    (F.Fmsy < 1 & B.Bmsy > 1) ~ "c"))
#Plot constants
y.max <- 1.5
x.max <- 10
all_missing <- stock_status %>%
  filter(is.na(B.Bmsy),is.na(F.Fmsy)) %>% 
  dplyr::select(Code, Council)
b_missing <- stock_status %>%
  filter(is.na(B.Bmsy), !is.na(F.Fmsy)) %>% 
  dplyr::select(Code, Council)
f_missing <- stock_status %>%
  filter(is.na(F.Fmsy), !is.na(B.Bmsy)) %>% 
  dplyr::select(Code, Council)
#A dataframe that defines custom legend for stocks with unknown status
all.df <- data.frame(text = all_missing$Code,
                    x = rep(x.max*0.9,length(all_missing$Code)),
                    #y = seq(1.45,1.05, length.out = 7))
                    y = seq(1.45,1.05, length.out = 8))
b.df <- data.frame(text = b_missing$Code,
                    x = rep(x.max*0.7,length(b_missing$Code)),
                    y = c(1.45,2.15))
f.df <- data.frame(text = f_missing$Code,
                    x = rep(x.max*0.5,length(f_missing$Code)),
                    y = seq(1.45,1.0, length.out = 8))

#Plotting code
ggplot(data = stock_status) +
  geom_vline(xintercept = 1, linetype = "dotted", color = "grey60")+
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "grey60")+
  geom_hline(yintercept = 1, linetype = "dashed", color = "grey60") +
  geom_point(aes(x = B.Bmsy,
                 y = F.Fmsy,
                 color = stock_status$score)) +
  geom_text_repel(aes(x = B.Bmsy, #geom_text_repel auto-jitters text around points
                      y = F.Fmsy,
                      label = Code,
                      color = stock_status$score), show.legend = FALSE,nudge_y = -0.01, nudge_x = 0.05) +
  ylim(0,y.max) +
  xlim(0,x.max*1.1) +
  geom_text(data = all.df, aes(x = x, y = y, label = text),show.legend = FALSE, size = 3)+
  geom_text(data = b.df, aes(x = x, y = y, label = text),show.legend = FALSE, size = 3)+
  geom_text(data = f.df, aes(x = x, y = y, label = text),show.legend = FALSE, size = 3)+
  scale_color_brewer(palette = "Dark2", #Change legend labels for clarity
                     breaks = stock_status$score) +
  annotate("rect", xmin = 0.924*x.max,
           xmax = 1.08*x.max,
           ymin = 0.645*y.max,
           ymax = 0.98*y.max,
           alpha = 0.01) +
  annotate("text", x = 9, y = 1.5, label = "F and B missing", fontface =2, size = 3)+
  annotate("rect",  
             xmin = 0.70*x.max,
             xmax = 0.85*x.max,
             ymin = 0.90*y.max,
             ymax = 1.8,
             alpha = 0.01) +
  annotate("text", x = 7, y = 1.5, label = "B missing", fontface =2, size = 3)+
  annotate("rect", xmin = 0.509*x.max,
           xmax = 0.681*x.max,
           ymin = 0.65*y.max,
           ymax = 0.98*y.max,
           alpha = 0.01) +
  annotate("text", x = 5, y = 1.5, label = "F missing", fontface =2, size = 3)+
  xlab(expression(~B/B[msy])) +
  ylab(expression(~F/F[msy])) +
  guides(color = FALSE) +
  theme_ts()

```


#### SOE 2020 ecodata source, New England

```{r, NEFMC2020plot}

stock_status <- ecodata::stock_status %>%
  mutate(Code = recode(Code, "Dogfish" = "Sp. Dogfish" )) %>% 
  spread(.,Var,Value) %>% 
  filter(Council %in% c("NEFMC","Both")) %>% 
  group_by(Stock) %>% 
  mutate(score = case_when(
    (B.Bmsy <0.5) ~"a",
    (F.Fmsy >1) ~ "a", 
    (F.Fmsy < 1 & B.Bmsy > 0.5 & B.Bmsy < 1) ~ "b",
    (F.Fmsy < 1 & B.Bmsy > 1) ~ "c"))
#Plot constants
y.max <- 1.5
x.max <- 10
all_missing <- stock_status %>%
  filter(is.na(B.Bmsy),is.na(F.Fmsy)) %>% 
  dplyr::select(Code, Council)
b_missing <- stock_status %>%
  filter(is.na(B.Bmsy), !is.na(F.Fmsy)) %>% 
  dplyr::select(Code, Council)
f_missing <- stock_status %>%
  filter(is.na(F.Fmsy), !is.na(B.Bmsy)) %>% 
  dplyr::select(Code, Council)
#A dataframe that defines custom legend for stocks with unknown status
all.df <- data.frame(text = all_missing$Code,
                    x = rep(x.max*0.9,length(all_missing$Code)),
                    y = seq(1.45,1.05, length.out = 7))
b.df <- data.frame(text = b_missing$Code,
                    x = rep(x.max*0.7,length(b_missing$Code)),
                    y = c(1.45,2.15))
f.df <- data.frame(text = f_missing$Code,
                    x = rep(x.max*0.5,length(f_missing$Code)),
                    y = seq(1.45,1.0, length.out = 8))

#Plotting code
ggplot(data = stock_status) +
  geom_vline(xintercept = 1, linetype = "dotted", color = "grey60")+
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "grey60")+
  geom_hline(yintercept = 1, linetype = "dashed", color = "grey60") +
  geom_point(aes(x = B.Bmsy,
                 y = F.Fmsy,
                 color = stock_status$score)) +
  geom_text_repel(aes(x = B.Bmsy, #geom_text_repel auto-jitters text around points
                      y = F.Fmsy,
                      label = Code,
                      color = stock_status$score), show.legend = FALSE,nudge_y = -0.01, nudge_x = 0.05) +
  ylim(0,y.max) +
  xlim(0,x.max*1.1) +
  geom_text(data = all.df, aes(x = x, y = y, label = text),show.legend = FALSE, size = 3)+
  geom_text(data = b.df, aes(x = x, y = y, label = text),show.legend = FALSE, size = 3)+
  geom_text(data = f.df, aes(x = x, y = y, label = text),show.legend = FALSE, size = 3)+
  scale_color_brewer(palette = "Dark2", #Change legend labels for clarity
                     breaks = stock_status$score) +
  annotate("rect", xmin = 0.924*x.max,
           xmax = 1.08*x.max,
           ymin = 0.645*y.max,
           ymax = 0.98*y.max,
           alpha = 0.01) +
  annotate("text", x = 9, y = 1.5, label = "F and B missing", fontface =2, size = 3)+
  annotate("rect",  
             xmin = 0.70*x.max,
             xmax = 0.85*x.max,
             ymin = 0.90*y.max,
             ymax = 1.8,
             alpha = 0.01) +
  annotate("text", x = 7, y = 1.5, label = "B missing", fontface =2, size = 3)+
  annotate("rect", xmin = 0.509*x.max,
           xmax = 0.681*x.max,
           ymin = 0.65*y.max,
           ymax = 0.98*y.max,
           alpha = 0.01) +
  annotate("text", x = 5, y = 1.5, label = "F missing", fontface =2, size = 3)+
  xlab(expression(~B/B[msy])) +
  ylab(expression(~F/F[msy])) +
  guides(color = FALSE) +
  theme_ts()


```


### {-}

### Issues

Mid-Atlantic: Mackerel B.Bmsy is different, lower in stockSMART.
New England: GOM Cod called F and B missing in stockSMART. Hakes are different. Wolffish missing from stockSMART?

### Next...

Then apply for assessments up to Dec 2020. Well, max assessment year is 2019, so I guess I still need to fill this in by hand!

Options:

1. Retain only the fields used by get_stocks.R so maybe I don't break ecodata.
1. Recreate the entire spreadsheet for all assessments (means they must all be in stocksmart).
