---
title: "2025 SOE stock status indicator"
author: "Sarah Gaichas"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

remotes::install_github("https://github.com/NOAA-EDAB/stocksmart")

# to have 2023's ecodata release for comparison
#remotes::install_github("noaa-edab/ecodata@v3.0",build_vignettes=TRUE)
# that didn't work I'm building it locally from my ESR/SOE2022/ecodata folder

library(here)
library(dplyr)
library(ggplot2)
library(tidyr)
library(DT)
library(stocksmart)
library(ggrepel)
library(ecodata)
library(ggiraph)

```

# Major Takeaway

Lets try stocksmart today, 31 Oct 2024, and see how much we have to update later if all the assessments aren't in.

As of 10 Dec 2024 stocksmart is not updated but data for spring 2024 MT assessments were sent via email from Kristan Blackhart to Scott. Dan Hennen created the files.

The file SISspringMT2024.zip has been included here and unzipped into the folder SISspringMT2024.

Stock specific status and output time series are in csv files which I will wrangle and inteleave with stocksmart data as below. 


## Pull from `stocksmart` for 2025 SOE reports and add PDB updates

Andy renamed the assessmentdata package [stocksmart](https://noaa-edab.github.io/stocksmart/) based on [Stock SMART](https://www.st.nmfs.noaa.gov/stocksmart?app=homepage). 

Two data frames are in the package, `stockAssessmentData` and `stockAssessmentSummary`.

In `stockAssessmentData` we have time series. Columns are `r names(stockAssessmentData)` and the reported metrics are `r unique(stockAssessmentData$Metric)`. 


```{r}

datatable(head(stockAssessmentData), rownames = FALSE)

```

In `stockAssessmentSummary` we have assessment metadata. Columns are `r (names(stockAssessmentSummary))`.

```{r}

datatable(head(stockAssessmentSummary), rownames = FALSE, options = list(scrollX = TRUE))

```

Build a stockSMART like object from PDB files from Spring 2024 MT assessments

```{r}

stocknames <- data.frame(
  `Stock / Entity Name` = c("Black sea bass - Mid-Atlantic Coast",
                            "Butterfish - Gulf of Maine / Cape Hatteras",
                            "Atlantic cod - Eastern Gulf of Maine",
                            "Atlantic cod - Georges Bank",
                            "Atlantic cod - Southern New England",
                            "Atlantic cod - Western Gulf of Maine",
                            "Tilefish - Mid-Atlantic Coast",
                            "Atlantic herring - Northwestern Atlantic Coast",
                            "Scup - Atlantic Coast"),
  ABBR = c("BSBUNIT",
           "BUTUNIT",
           "CODEGOM",
           "CODGB",
           "CODSNE",
           "CODWGOM",
           "GTFUNIT",
           "HERUNIT",
           "SCUNIT")
)

getstatus <- function(fname){
   #need name of stock from filename then read in other data
  statdat <- read.csv(fname, header = FALSE)
  statdat[statdat == "?"] <- NA
  statdat <- statdat |>
    dplyr::filter(V2!="") |> 
    tidyr::pivot_wider(names_from = V1, values_from = V2)
  
}

# get list of filenames


# purrr map the list of filenames with getstatus into a dataframe


# rbind them with current stocksmart?


```


Build ecodata input spreadsheet from `stockAssessmentSummary` and use the ecodata code to make the dataset for plotting:

```{r make-2022assess}

assess2024 <- stocksmart::stockAssessmentSummary |>
  dplyr::filter(`Science Center` == "NEFSC") |>
  dplyr::select(c(`Stock Name`, Jurisdiction, FMP, `Science Center`, 
                  `Stock Area`, `Assessment Year`, `Last Data Year`,
                  `F Year`, `Estimated F`, Flimit, Fmsy, `F/Flimit`, 
                  `F/Fmsy`, Ftarget, `F/Ftarget`, `B Year`, `Estimated B`,
                  `B Unit`, Blimit, Bmsy, `B/Blimit`, `B/Bmsy`)) |>
  dplyr::arrange(Jurisdiction, `Stock Name`, FMP, `Assessment Year`) |>
  dplyr::rename(Entity.Name = `Stock Name`,
                Assessment.Year = `Assessment Year`,
                F.Fmsy = `F/Fmsy`,
                B.Bmsy = `B/Bmsy`,
                Estimated.F = `Estimated F`,
                Estimated.B = `Estimated B`)

write.csv(assess2024, here("assess.csv"))

# from get_stocks.R, ecodata 2020
  #assess <- read.csv(file.path(data.dir, "2019assess.csv"))
  assess <- assess2024
  #decode <- read.csv(file.path(data.dir, "2019decoder.csv"))
  decode <- read.csv(here("2020decoder.csv"))
  
write.csv(decode, here("decoder.csv"))

stock_status_stockSMART <-
    assess %>%
    dplyr::group_by(Entity.Name) %>%
    dplyr::filter(Assessment.Year == max(Assessment.Year)) %>%
  #Find last year assessment occurred for each stock
    dplyr::ungroup() %>%
    dplyr::left_join(.,decode, by = "Entity.Name") %>% #Join in list of managed species
    dplyr::select(Entity.Name, Assessment.Year, F.Fmsy, B.Bmsy, Council, Code) %>%
  #select column variables to keep
    dplyr::mutate(id = 1:length(Entity.Name)) %>%
    tidyr::gather(.,Var, Value,-id,-Entity.Name,-Assessment.Year,-Council,-Code) %>%
  #wide to long
    dplyr::select(-id) %>%
    dplyr::rename(`Last assessment` = Assessment.Year,
                  Stock = Entity.Name) %>% #rename variables for clarity
    dplyr::mutate(Units = "unitless") #%>%
    #dplyr::mutate(Value = replace(Value, which(Code == "N Windowpane" & Var == "F.Fmsy"), NA))



```

Then test to see if we see the updates relative to SOE 2023 (updates through 2022) ecodata. I'm leaving out all the plot annotations for unknown status. 

### Comparisons {.tabset}

#### StockSMART Oct 2024 source, Mid-Atlantic

```{r, MAFMCplot}

stock_status <- 
  stock_status_stockSMART %>%
  mutate(Code = recode(Code, "Dogfish" = "Sp. Dogfish" )) %>% 
  spread(.,Var,Value) %>% 
  filter(Council %in% c("MAFMC","Both")) %>% 
  group_by(Stock) %>% 
  mutate(score = case_when(
    (B.Bmsy <0.5) ~"a",
    (F.Fmsy >1) ~ "a", 
    (F.Fmsy < 1 & B.Bmsy > 0.5 & B.Bmsy < 1) ~ "b",
    (F.Fmsy < 1 & B.Bmsy > 1) ~ "c"))
#Plot constants
y.max <- 2.1 #2.0 mackerel cut off F/Fmsy is 2.08
x.max <- 2.6
#A dataframe that defines custom legend for stocks with unknown status
# unknown <- data.frame(text = c("Unknown Status", "Longfin Squid",
#                               "Shortfin Squid", "N. Goosefish", "S. Goosefish"),
#                     x = rep(0.9*x.max,5), y = seq(0.93*y.max,1.5,-.1))

# Custom Color
custom_color<- c("#56B4E9", "#009E73", "#0072B2")
#Plotting code
ggplot(data = stock_status) +
  geom_vline(xintercept = 1, linetype = "dotted")+
  geom_vline(xintercept = 0.5, linetype = "dashed")+
  geom_hline(yintercept = 1, linetype = "dashed") +
  geom_point(aes(x = B.Bmsy,
                 y = F.Fmsy,
                 shape = Council,
                 color = score)) +
  geom_text_repel(aes(x = B.Bmsy, #geom_text_repel auto-jitters text around points
                      y = F.Fmsy,
                      label = Code, 
                      color = score), 
                  show.legend = FALSE, nudge_y = -0.01, nudge_x = 0.05) +
  scale_color_brewer(palette = "Dark2",
                     breaks = stock_status$score) +
  ylim(0,y.max) +
  xlim(0,x.max) +
  # geom_text(data = unknown, aes(x = x, y = y, label = text), #Custom legend for unknown stock status
  #           size = c(4.75,rep(4,4))) +
  # annotate("rect", xmin = 0.8*x.max,
  #          xmax = x.max,
  #          ymin = 0.65*y.max,
  #          ymax = 0.90*y.max,
  #          alpha = 0.1) +
  xlab(expression(~B/B[msy])) +
  ylab(expression(~F/F[msy])) +
  guides(color = FALSE) +
  theme_ts()
```

#### SOE 2024 ecodata source, Mid-Atlantic

```{r, MAFMC2022plot}
#Get data, spread for plotting, and filter
stock_status <- ecodata::stock_status %>%
  dplyr::mutate(Code = recode(Code, "Dogfish" = "Sp. Dogfish" ), 
                Code = recode(Code, "Mackerel" = "At. Mackerel")) %>% 
  tidyr::spread(.,Var,Value) %>% 
  dplyr::filter(Council %in% c("MAFMC","Both")) %>% 
  dplyr::group_by(Stock) %>% 
  dplyr::mutate(score = case_when(
    (B.Bmsy <0.5) ~"a",
    (F.Fmsy >1) ~ "a", 
    (F.Fmsy < 1 & B.Bmsy > 0.5 & B.Bmsy < 1) ~ "b",
    (F.Fmsy < 1 & B.Bmsy > 1) ~ "c"))
#Plot constants
y.max <- 2.1 #1.75 mackerel cut off F/Fmsy is 1.8
x.max <- 2.6
#A dataframe that defines custom legend for stocks with unknown status
unknown <- data.frame(text = c("Unknown Status", "Longfin Squid",
                              "Shortfin Squid", "N. Goosefish", "S. Goosefish", "Blueline Tilefish", "Chub Mackerel"),
                    x = rep(0.9*x.max,7), y = seq(0.88*y.max,1.2,-0.1))

# Custom Color
custom_color<- c("#56B4E9", "#009E73", "#0072B2")
#Plotting code
ggplot2::ggplot(data = stock_status) +
  ggplot2::geom_vline(xintercept = 1, linetype = "dotted")+
  ggplot2::geom_vline(xintercept = 0.5, linetype = "dashed")+
  ggplot2::geom_hline(yintercept = 1, linetype = "dashed") +
  ggplot2::geom_point(aes(x = B.Bmsy,
                 y = F.Fmsy,
                 shape = Council,
                 color = score)) +
  ggrepel::geom_text_repel(aes(x = B.Bmsy, #geom_text_repel auto-jitters text around points
                      y = F.Fmsy,
                      label = Code, 
                      color = score), 
                  show.legend = FALSE, nudge_y = -0.01, nudge_x = 0.05) +
  ggplot2::scale_color_brewer(palette = "Dark2",
                     breaks = stock_status$score) +
  ggplot2::ylim(0,y.max) +
  ggplot2::xlim(0,x.max) +
  ggplot2::geom_text(data = unknown, aes(x = x-0.5, y = y+0.2, label = text), #Custom legend for unknown stock status
            size = c(4.75,rep(4,6))) +
  #ggplot2::annotate("rect", xmin = 0.8*x.max,
  #         xmax = x.max
  #         ymin = 0.65*y.max,
  #         ymax = 0.90*y.max,
  #         alpha = 0.1) +
  ggplot2::xlab(expression(~B/B[msy])) +
  ggplot2::ylab(expression(~F/F[msy])) +
  ggplot2::guides(color = FALSE) +
  ecodata::theme_ts()
```


#### StockSMART Oct 2024 source, New England

```{r, NEFMCtest}

stock_status <- stock_status_stockSMART %>%
  mutate(Code = recode(Code, "Dogfish" = "Sp. Dogfish" )) %>% 
  spread(.,Var,Value) %>% 
  filter(Council %in% c("NEFMC","Both")) %>% 
  group_by(Stock) %>% 
  mutate(score = case_when(
    (B.Bmsy <0.5) ~"a",
    (F.Fmsy >1) ~ "a", 
    (F.Fmsy < 1 & B.Bmsy > 0.5 & B.Bmsy < 1) ~ "b",
    (F.Fmsy < 1 & B.Bmsy > 1) ~ "c"))
#Plot constants
y.max <- 1.5
x.max <- 10
all_missing <- stock_status %>%
  filter(is.na(B.Bmsy),is.na(F.Fmsy)) %>% 
  dplyr::select(Code, Council)
b_missing <- stock_status %>%
  filter(is.na(B.Bmsy), !is.na(F.Fmsy)) %>% 
  dplyr::select(Code, Council)
f_missing <- stock_status %>%
  filter(is.na(F.Fmsy), !is.na(B.Bmsy)) %>% 
  dplyr::select(Code, Council)
#A dataframe that defines custom legend for stocks with unknown status
# all.df <- data.frame(text = all_missing$Code,
#                     x = rep(x.max*0.9,length(all_missing$Code)),
#                     #y = seq(1.45,1.05, length.out = 7))
#                     y = seq(1.45,1.05, length.out = length(all_missing$Code)))
# b.df <- data.frame(text = b_missing$Code,
#                     x = rep(x.max*0.7,length(b_missing$Code)),
#                     y = c(1.45,2.15, length.out = length(b_missing$Code)))
# f.df <- data.frame(text = f_missing$Code,
#                     x = rep(x.max*0.5,length(f_missing$Code)),
#                     y = seq(1.45,1.0, length.out = length(f_missing$Code)))

# Custom Color
custom_color<- c("#56B4E9", "#009E73", "#0072B2")

#Plotting code
ggplot(data = stock_status) +
  geom_vline(xintercept = 1, linetype = "dotted", color = "grey60")+
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "grey60")+
  geom_hline(yintercept = 1, linetype = "dashed", color = "grey60") +
  geom_point(aes(x = B.Bmsy,
                 y = F.Fmsy,
                 color = stock_status$score)) +
  geom_text_repel(aes(x = B.Bmsy, #geom_text_repel auto-jitters text around points
                      y = F.Fmsy,
                      label = Code,
                      color = stock_status$score), show.legend = FALSE,nudge_y = -0.01, nudge_x = 0.05) +
  ylim(0,y.max) +
  xlim(0,x.max*1.1) +
  # geom_text(data = all.df, aes(x = x, y = y, label = text),show.legend = FALSE, size = 3)+
  # geom_text(data = b.df, aes(x = x, y = y, label = text),show.legend = FALSE, size = 3)+
  # geom_text(data = f.df, aes(x = x, y = y, label = text),show.legend = FALSE, size = 3)+
  # scale_color_brewer(palette = "Dark2", #Change legend labels for clarity
  #                    breaks = stock_status$score) +
  # annotate("rect", xmin = 0.924*x.max,
  #          xmax = 1.08*x.max,
  #          ymin = 0.645*y.max,
  #          ymax = 0.98*y.max,
  #          alpha = 0.01) +
  # annotate("text", x = 9, y = 1.5, label = "F and B missing", fontface =2, size = 3)+
  # annotate("rect",  
  #            xmin = 0.70*x.max,
  #            xmax = 0.85*x.max,
  #            ymin = 0.90*y.max,
  #            ymax = 1.8,
  #            alpha = 0.01) +
  # annotate("text", x = 7, y = 1.5, label = "B missing", fontface =2, size = 3)+
  # annotate("rect", xmin = 0.509*x.max,
  #          xmax = 0.681*x.max,
  #          ymin = 0.65*y.max,
  #          ymax = 0.98*y.max,
  #          alpha = 0.01) +
  # annotate("text", x = 5, y = 1.5, label = "F missing", fontface =2, size = 3)+
  xlab(expression(~B/B[msy])) +
  ylab(expression(~F/F[msy])) +
  guides(color = FALSE) +
  theme_ts()

```


#### SOE 2024 ecodata source, New England

```{r, NEFMC2022plot}
#Get data, spread for plotting, and filter
stock_status <- ecodata::stock_status %>%
  dplyr::mutate(Code = dplyr::recode(Code, "Dogfish" = "Sp. Dogfish" )) %>% 
  tidyr::spread(.,Var,Value) %>% 
  dplyr::filter(Council %in% c("NEFMC","Both")) %>% 
  dplyr::group_by(Stock) %>% 
  dplyr::mutate(score = case_when(
    (B.Bmsy <0.5) ~"a",
    (F.Fmsy >1) ~ "a", 
    (F.Fmsy < 1 & B.Bmsy > 0.5 & B.Bmsy < 1) ~ "b",
    (F.Fmsy < 1 & B.Bmsy > 1) ~ "c"))
#Plot constants
y.max <- 1.5
x.max <- 10
all_missing <- stock_status %>%
  dplyr::filter(is.na(B.Bmsy),is.na(F.Fmsy)) %>% 
  dplyr::select(Code, Council)
b_missing <- stock_status %>%
  dplyr::filter(is.na(B.Bmsy), !is.na(F.Fmsy)) %>% 
  dplyr::select(Code, Council)
f_missing <- stock_status %>%
  dplyr::filter(is.na(F.Fmsy), !is.na(B.Bmsy)) %>% 
  dplyr::select(Code, Council)
#A dataframe that defines custom legend for stocks with unknown status
all.df <- data.frame(text = all_missing$Code,
                    x = rep(x.max*0.9,length(all_missing$Code)),
                    y = seq(1.45,0.7, length.out = length(all_missing$Code)))
b.df <- data.frame(text = b_missing$Code,
                    x = rep(x.max*0.7,length(b_missing$Code)),
                    y = seq(1.45,1.30, length.out = length(b_missing$Code)))
f.df <- data.frame(text = f_missing$Code,
                    x = rep(x.max*0.5,length(f_missing$Code)),
                    y = seq(1.45,1, length.out = length(f_missing$Code)))

#Plotting code
ggplot2::ggplot(data = stock_status) +
  ggplot2::geom_vline(xintercept = 1, linetype = "dotted", color = "grey60")+
  ggplot2::geom_vline(xintercept = 0.5, linetype = "dashed", color = "grey60")+
  ggplot2::geom_hline(yintercept = 1, linetype = "dashed", color = "grey60") +
  ggplot2::geom_point(aes(x = B.Bmsy,
                 y = F.Fmsy,
                 color = stock_status$score)) +
  ggrepel::geom_text_repel(aes(x = B.Bmsy, #geom_text_repel auto-jitters text around points
                      y = F.Fmsy,
                      label = Code,
                      color = stock_status$score), show.legend = FALSE,nudge_y = -0.01, nudge_x = 0.05) +
  ggplot2::ylim(0,y.max) +
  ggplot2::xlim(0,x.max*1.1) +
  ggplot2::geom_text(data = all.df, aes(x = x, y = y, label = text),show.legend = FALSE, size = 3)+
  ggplot2::geom_text(data = b.df, aes(x = x, y = y, label = text),show.legend = FALSE, size = 3)+
  ggplot2::geom_text(data = f.df, aes(x = x, y = y, label = text),show.legend = FALSE, size = 3)+
  ggplot2::scale_color_brewer(palette = "Dark2", #Change legend labels for clarity
                     breaks = stock_status$score) +
  ggplot2::annotate("rect", xmin = 0.924*x.max,
           xmax = 1.08*x.max,
           ymin = 0.645*y.max,
           ymax = 0.98*y.max,
           alpha = 0.01) +
  ggplot2::annotate("text", x = 9, y = 1.5, label = "F and B missing", fontface =2, size = 3)+
  ggplot2::annotate("rect",  
             xmin = 0.70*x.max,
             xmax = 0.85*x.max,
             ymin = 0.30*y.max,
             ymax = 0.98*y.max,
             alpha = 0.01) +
  ggplot2::annotate("text", x = 7, y = 1.5, label = "B missing", fontface =2, size = 3)+
  ggplot2::annotate("rect", xmin = 0.509*x.max,
           xmax = 0.681*x.max,
           ymin = 0.65*y.max,
           ymax = 0.98*y.max,
           alpha = 0.01) +
  ggplot2::annotate("text", x = 5, y = 1.5, label = "F missing", fontface =2, size = 3)+
  ggplot2::xlab(expression(~B/B[msy])) +
  ggplot2::ylab(expression(~F/F[msy])) +
  ggplot2::guides(color = FALSE) +
  ecodata::theme_ts()


```


### {-}

### Issues

None noted, stockSMART and `stocksmart` have been updated as of December 13 2023. 




